### 2020-03-11 Common/chapters/transfer_learning.tex
Julien and Johan: proofread til the end

First, a general definition of transfer learning is given. Its utility in this specific application is discussed as it is useful but not in any case. Second, the different steps involved in the process are described in order to get an overview of the whole transfer learning pipeline that was applied. Then, a section focuses on the processing of new datasets that are used in the final experiment. Finally, the latter and its results are discussed.

@@

\section{Process overview}
\label{sec:process_overview}
\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}The approach consists in five different steps. Figure \ref{fig:tl_overview} gives a visual representation of the following explanations:\\
\begin{enumerate}
\item First of all, similar datasets are collected and processed in the same way. These datasets need to share common features. For example, DWI images (MRI sequence) and CT scans share the same color scheme and look alike (the same kind of tissue is visible, etc.).
%On the contrary, mammographies are completely different in that they the tissue is less visible FIGURE.
\item Second, the convolutional neural network is built and split into two parts. It is still a single network but the first layers are treated as the feature extractor. As the name suggests, the role of these layers is to extract visual features characterizing what a cancerous tumor looks like. On the other hand, the last layers are considered as decision maker. They will gather the features collected by the feature extractor and process them to make a decision on the class of the input sample.
\item Third, the model is trained on the target dataset and the model whose performance is the best is saved. The target dataset is the dataset which one wants to improve the performance of the model on. This first training establishes a baseline of performance and of features to learn further on thanks to the other datasets.
\item Fourth, the core of transfer learning process starts. For each non-target dataset, two different trainings take place:

\begin{enumerate}
	\item The model saved at the end of the previous training is loaded. Its last layers (the decision maker) are completely reset, i.e. the weights are reset to another initialization, as if no training occured for these layers. Afterwards, the first layers (the features extractor) are frozen and the model is trained. Since the first layers have been frozen, their weights will remain the same. Only the weights of the last layers will change. This part of the transfer learning allows the model to make use of the previously learned features while adapting to the new body part by training the decision maker. 
	\item Once the training with frozen layers is over, the best model is loaded, the frozen layers are unfrozen, and a normal training is performed on the same dataset. Thanks to this, the model adapts the previously learned features and tries to learn new features which could be useful for the classification of the target dataset. 
\end{enumerate}
\item Fifth, the previous step is also applied to the target dataset. The only difference lies in the frozen part. Instead of dropping the old last layers and initialize its weights randomly, the decision maker of the best model found during the training on the first dataset is attached to the model. These final two trainings aim at keeping the interesting features from the other datasets while learning specializing the model in the target dataset again. Hopefully, the new weight initialization resulting from the transfer learning will allow to reach better metric values than before the transfer learning.

\end{enumerate}
\begin{figure}[!h]
\centering
\noindent
\makebox[\textwidth]{
\includegraphics[width=1\textwidth, keepaspectratio=true]{./figures/tl_overview.png}
}
\caption{Transfer learning - Overview}
\label{fig:tl_overview}
\end{figure}

@@

The overall process is the same as the one described in section \ref{sec:numpyToAugmentedStacked}, apart from a few details. First, a T2-weighted image was still used to align and resize the corresponding DWI image, but was not exported at the end, resulting in a single-channel DWI image (NumPy array) instead of stacked images. As DWI images are not of the same size, the fact of using the corresonding T2 images to resize them was the easiest way to make the new implementation go without a hitch.

@@

\subsection{LungCTChallenge}
\label{sec:lungCTChallenge}
\subsubsection{Dataset description}
\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}Lung CT Challenge ~\cite{12, 36, 38} is composed of two different subdatasets of CT scans: one is called "calibration set" (10 patients) and the other "test set" (60 patients). Each patient can have one or multiple findings. Since labels were provided for both sets and the amount of data is fairly low, they were merged and used as a global training set. For each finding, an abnormal mass is visible in the lungs, which means that the classification is going to differentiate malignant from benign tumors.

Regarding labelling, two Excel files, \textit{TestSet\_NoduleData\_PublicRelease\_wTruth} and \textit{CalibrationSet\_NoduleData}, contain labels for these images. In order to facilitate the handling of information in the code, two CSV files were manually created: \mbox{\textit{TestSet.csv}} and \textit{CalibrationSet.csv}.
Contrary to PROSTATEx, more than two labels were used in this dataset. Both "malignant" and "Primary lung cancer" were considered as positive, whereas "benign" and "Benign nodule" were treated as negative. A third label called "Suspicious malignant nodule" appeared two times. Since the diagnosis was not clearly defined for those images, they were not included in the training data to avoid any noise.

\subsubsection{From DICOM to augmented NumPy arrays}
\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}As PROSTATEx, the processing was split into two parts: preprocessing (DICOM to NumPy arrays) and augmentation (NumPy arrays to cropped augmented NumPy arrays.) Algorithm \ref{alg:LungCTChallenge_preprocessing} shows the various preprocessing steps.

The augmentation process is close to the one applied to PROSTATEx. An advantage Lung CT Challenge has is that all images have the same resolution (512x512 pixels), making the alignment process used for PROSTATEx useless. Nevertheless, lung tumors are particular. What differentiates them from the others is their look. In fact, lung tumors usually look like spider webs with a central nodule and branches coming out of it, whereas brain and prostate tumors tend to have well-defined boundaries. Therefore, a larger patch size was chosen in order to keep the extra information provided by the branches in the peripheral tissue. Furthermore, a perfect cropping method would only crop the interesting part of the tissue and a margin around it. However, tumor size varies a lot. It can either be extremely small (a dot) or large (massive nodule with long branches). Consequently, the selected patch size makes the largest tumors fit perfectly in the cropped area, inducing more tissue for smaller tumors. Naturally, there were few outliers whose size was a lot larger than the chosen patch size. If the patch size was set accordingly, entire lungs were sometimes visible in other patients' case. For this reason, these large tumors were treated as outliers. The chosen dimensions are 1/6 of the image width and height (i.e. 85x85px since the image size is 512x512px). These patches were then resized to 65x65px thanks to a bicubic interpolation, before being exported as NumPy arrays.

@@

\subsection{Kaggle Brain}
\label{sec:kaggleBrain}
\subsubsection{Dataset description}
\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}This dataset is publicly available on Kaggle~\cite{51}. All images come from single-sequence MRIs. It is composed of two folders: one containing images with cancerous tumors and another one containing images either without tumor or with a non-cancerous tumor. Images whose quality was extremely low (resolution, picture of a screen, artefacts, etc.) were manually removed. In other cases, the image quality was good but disturbing objects such as left/right markers, arrows pointing the tumor, etc. were visible. These objects were manually removed using the object removal tool in Photoshop or by cropping the image if the disturbing object was far from the skull. File names and file formats were heterogeneous. At the end of this process, the dataset was composed of 96 benign lesions and 133 malignant ones.

@@

The first step was to standardize the file names and formats. The chosen convention is "$[yes|no]{index}.jpg$". The vast majority of the files were in the JPG file format, so the few PNG images were converted. Then, a CSV file containing the following columns was created: "PatientID", "Nodule Center Position", "fid", "Diagnosis". The "PatientID" is the filename without extension that was set during the first step (for example "$yes99$" or "$no3$"). The "Nodule Center Position" represents the pixel coordinates of the lesion. Fortunately, brain tumors are clearly visible on brain MRIs, which eased the process. To find the coordinates, every single image was opened at its original size in an image visualization software. A rectangle selection tool was used to draw a rectangle from the top left corner of the image (because the top left pixel corresponds to the pixel at the coordinates $(0,0)$ in NumPy arrays) to the center of the lesion. The pixel width and height of this rectangle gave the "i,j" coordinates of the lesion. These coordinates were reported manually. Then, the "fid" corresponds to the Finding ID. This parameter would be more pertinent if the same "PatientID" could have multiple findings, i.e. if the same patient could have multiple tumors. Since this information is made use of during the image processing of the other datasets, it was also added here. In this particular case, the "fid" was always set to "1". Finally, the "Diagnosis" was defined as "False" or "True" according to the folder which the image was in. This allows to simply use the same processing techniques as it is done for the other datasets.

@@

Similarly to Lung CT Challenge, tumor sizes vary substantially. The same methodology was applied regarding the patch size, i.e. choosing a size which fits the majority of large tumors perfectly. Even though the images all show the head under the same perspective (same height/width ratio, visible skull, black margin around the skull), images were probably collected from different sources, which explains the heterogeneity of resolutions. As a consequence, the patch size could not be a fixed number of pixels for all images because the cropped area would be the entire head on one image and a small portion of the tumor on another. The trick consisted in finding the horizontal and vertical portions of the image occupied by the tumor. In fact, cropping an area centered on the lesion whose height was $\frac{2}{7}$ of the image and whose width was $\frac{1}{4}$ gave good results. Finally, images were also resized to 65x65px using a bicubic interpolation before being saved as NumPy arrays.

\subsection{Image cropping}
\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}Multiple ways of cropping images were tested out. The differences lay in how much tissue was kept around the lesion. In some cases, more context around the lesion allows to capture details about the connections between the lesion and blood vessels for example. Moreover, the abrupt borders inducing a clear change of colors can help detect tumors. Experienced showed that more context helps when the images are single-sequence. On the other hand, cropping right around the lesion or even in the lesion gave better results for multi-sequence data, i.e. when T2-weighted, DWI and ADC images (MRIs) were simultaneously used to perform the tumor detection and classification. In the latter case, the lesion usually appears distinctively on the three sequences, and this impacts the decision a lot. On single-sequence images, there is a clear lack of information which makes it difficult to differentiate a tumor from another tissue without much context.

\subsection{Verification}
\subsubsection{Visual checking}
\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}Regarding the Kaggle brain dataset, the data generated was paid close attention to since the ground truth CSV file was created by hand (see Section \ref{sec:braingroundtruth}). In fact, each generated image was taken a close look at, which allowed to spot 3 mistakes over the whole dataset. They concerned typos in the "Nodule Center Position", which created pitch black images since the cropped area was outside.
Like PROSTATEx (see section \ref{sec:dataProcessingVerification}), control images with red dots were also generated for the other datasets.

\section{Transfer learning implementation}
\subsection{Layer freezing}
\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}After each dataset switch, the model is first trained with freezed layers. The freezed layers are those considered as feature extractor, i.e. the first layers of the model. The exact point where the model is split is arbitrary, but is usually located after the convolutions. To perform this freezing, multiple steps are required:
\begin{enumerate}
	\item Reset the last layers. To do so, an instance of the model is created by loading the state dictionary of the previous one, before replacing the module containing the last layers with a new instance. 
	\item Iterate over the batch normalization layers and freeze them. Step 3 is not sufficient to freeze batch normalization layers. This can be achieved by calling module.eval() on each module containing such layers (after the model.train() call). 
	\item Only pass the parameters of these last layers to the optimizer.
\end{enumerate}

@@

\subsection{Conditions to choose the best model}
\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}At each step of the transfer learning, the best model was saved in order to be used during the next transfer learning step. After each epoch, the current model was compared to the previous best model. If it was better, the current model became the best model. A model was considered better if its AUC was higher than the best model's AUC, and if its accuracy was higher than the best model's accuracy. As the goal was to optimize the AUC, the first condition makes immediate sense. The second condition was put in place in order to prevent a sudden spike in AUC to set an unreacheable value. This scenario did not occur during this experiment but during previous ones with random initialization. The AUC at early epochs could be really high, saving an untrained model the best model.

@@

\subsubsection{Automated transfer learning}
\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin}}The first Python script runs the entire transfer learning pipeline by itself. Multiple command line arguments are required in order to load the right datasets and the right model, to define the hyperparameters, etc. Table \ref{table:transfer_learning_options} gives an exhaustive list of the script options. The particularity lies in the learning rate, number of epochs, batch size and dropout options. In fact, the pipeline consists in seven different phases of training: dataset 1 (full model), dataset 2 (frozen model), dataset 2 (full), dataset 3 (frozen), dataset 3 (full), dataset 4 (frozen) and dataset 4 (full). In such a case, each phase will probably use different hyperparameters. Therefore, these command line options need to specify seven different parameters. For example, a valid value for the "-{}-lr" option could be "1e-8,1e-5,1e-6,1e-6,1e-5,1e-7,1e-8".

\subsubsection{Manual transfer learning}
\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}The second Python script allows to perform each transfer learning step by hand. Thanks to this, a good combination of hyperparameters can be searched for without having to run the whole pipeline. The command line arguments slighty differ from the automated version. Table \ref{table:manual_transfer_learning_options} lists them. First of all, a single training set and a single validation set have to be specificied. In addition to this, a second dataset is required: the reference/target one, which does not change throughout the process. This dataset is useful for the global performance graphs (see Section \ref{sec:global_performance}). Another parameter used for the global performance graphs is the '-{}-last{\_}layer'. Its role is limited to this, unless '-{}-{}attach{\_}DM' is set to True. In this case, the decision maker loaded as '-{}-last{\_}layer' is attached to the current model, provided the current step is a frozen one. This option is used at the last frozen step (target dataset), as this decision maker comes from the first training (target dataset as well). Doing this instead of a using a new random decision maker (as it is the case for the non-target datasets) allows to start with a skilled decision maker and optimize it according to the features learned during the transfer learning.

Furthermore, arguments which were comma-separated values become single values, since this file runs a single transfer learning step instead of seven. Also, an output name for Tensorboard needs to be specified. In the automated version, this output name is automatically handled by the code itself. The latter allows to organize the various graphs generated by Tensorboard within the Tensorboard interface. If it is used in a correct way, Tensorboard can then be run on the root directory containing all graphs (which may be in subdirectories).

@@

\textbf{-{}-cv} & \begin{tabular}[c]{@{}l@{}}Whether to use cross-validation or not\\{[}'True', 'False'{]}\end{tabular} & \begin{tabular}[c]{@{}l@{}}True\end{tabular}             

@@

\subsection{Visualization of the impact of the various datasets on the target task}
\label{sec:global_performance}
\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}The so-called global performance graphs are a visual representation of the impact of each part of the transfer learning and each dataset on the target task. On paper, each step of the transfer learning is supposed to make the model learn new features that are useful for the first dataset (the target dataset). As described in section \ref{sec:process_overview}, the model can be seen as a feature extractor and a decision maker. Hence, to see how much each step was actually helping the process, the decision maker resulting from the best model of the first step of the transfer learning, i.e. the training on the first/target dataset, was saved. As this decision maker is saved after the training on the target dataset, it is specialized in the classification of the latter. In other words, if the target dataset is a prostate dataset, it will classify prostate tumors with a relatively high precision.

Then, after each epoch of every other transfer learning step, the saved decision maker was attached to the current feature extractor. This temporary model was then tested on the target training and validation sets, which resulted in different training and validation metrics.

At the end of the transfer learning pipeline, a graph was generated for each metric. These curves show the evolution of the latter across the entire process, which allows to determine the impact of each non-target dataset on the target task. If a metric is increasing, the features learned thanks to the corresponding dataset are useful for the target one. In this case, the graphs show how much the features learned from each dataset help classify prostate lesions.

@@

\subsection{Experimental setup}
\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin}}The whole transfer learning consists in multiple steps. In order to maximize the performance during the whole process, the hyperparameters were manually tuned at each step, following a grid search approach to find the best ones. Table \ref{table:tl_hyperparameters_at_each_step} presents the best hyperparameters used at each step of the experiment. Note that, as it was the case in the paper experiment reproduction, the first model was also initialized by loading the best model among 200 initializations. The model was implemented with PyTorch using the Adam optimizer to update its weights. The complete succession of the process took approximately 60 hours on an Nvidia GeForce RTX 2080 Ti. 
%Tesla v100 XSM2 graphics card.

>>>>

First, a general definition of transfer learning is given. Its utility in this specific application is discussed as it can be useful but not in any case. Second, the different steps involved in the process are described in order to get an overview of the whole transfer learning pipeline that was applied. Then, a section focuses on the processing of new datasets that are used in the final experiment. Finally, the latter and its results are discussed.

@@

\section{Process overview}
\label{sec:process_overview}
\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}The approach consists in five different steps. Figure \ref{fig:tl_overview} gives a visual representation of the following explanations:\\
\begin{enumerate}
\item First of all, similar datasets are collected and processed in the same way. These datasets need to share common features. For example, DWI images (MRI sequence) and CT scans share the same color scheme and look alike (the same kind of tissue is visible, etc.).
%On the contrary, mammographies are completely different in that they the tissue is less visible FIGURE.
\item Second, the convolutional neural network is built and split into two parts. It is still a single network but the first layers are treated as a feature extractor. As the name suggests, the role of these layers is to extract visual features characterizing what a cancerous tumor looks like. On the other hand, the last layers are considered as a decision maker. They gather the features collected by the feature extractor and process them to make a decision on the class of the input sample.
\item Third, the model is trained on the target dataset and the model whose performance is the best is saved. The target dataset is the dataset which one wants to improve the performance of the model on. This first training establishes a baseline of performance and a baseline of features which will be improved further on.
\item Fourth, the core of the transfer learning process starts. For each non-target dataset, two different trainings take place:

\begin{enumerate}
	\item The model saved at the end of the previous training is loaded. Its last layers (the decision maker) are completely reset, i.e. the weights are reset to another initialization, as if no training occured for these layers. Afterwards, the first layers (the feature extractor) are frozen and the model is trained. Since the first layers have been frozen, their weights remain the same. Only the weights of the last layers change during this training. This part of the transfer learning allows the model to make use of the previously learned features while adapting to the new body part by training the decision maker. 
	\item Once the training with frozen layers is over, the best model is loaded, the frozen layers are unfrozen, and a normal training is performed on the same dataset. Thanks to this, the model adapts the previously learned features and tries to learn new features which could be useful for the classification of the target dataset. 
\end{enumerate}
\item Fifth, the previous step is also applied to the target dataset. The only difference lies in the frozen part. Instead of dropping the old last layers and initialize their weights randomly, the decision maker of the best model found during the training on the first dataset is attached to the model. These final two trainings aim at keeping the interesting features from the other datasets while learning specializing the model in the target dataset again. Hopefully, the new weight initialization resulting from the transfer learning will allow to reach better metric values than before the transfer learning.

\end{enumerate}
\begin{figure}[!h]
\centering
\noindent
\makebox[\textwidth]{
\includegraphics[width=1.1\textwidth, keepaspectratio=true]{./figures/tl_overview.png}
}
\caption{Transfer learning - Overview}
\label{fig:tl_overview}
\end{figure}

@@

The overall process is the same as the one described in Section \ref{sec:numpyToAugmentedStacked}, apart from a few details. First, a T2-weighted image was still used to align and resize the corresponding DWI image, but was not exported at the end, resulting in a single-channel DWI image (NumPy array) instead of a stacked image. As DWI images are not of the same size, the fact of using the corresponding T2 images to resize them was the easiest way to make the new implementation go without a hitch.

@@

\subsection{LungCTChallenge}
\label{sec:lungCTChallenge}
\subsubsection{Dataset description}
\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}Lung CT Challenge ~\cite{12, 36, 38} is composed of two different subdatasets of CT scans: one is called "calibration set" (10 patients) and the other "test set" (60 patients). Each patient can have one or multiple findings. As labels were provided for both sets and the amount of data is fairly low, they were merged and used as a global training set. For each finding, an abnormal mass is visible in the lungs, which means that the classification is going to differentiate malignant from benign tumors.

Regarding labelling, two Excel files, \textit{TestSet\_NoduleData\_PublicRelease\_wTruth} and \textit{CalibrationSet\_NoduleData}, contain the corresponding labels for these images. In order to facilitate the handling of information in the code, two CSV files were manually created: \mbox{\textit{TestSet.csv}} and \textit{CalibrationSet.csv}.
Contrary to PROSTATEx, more than two labels were used in this dataset. Both "malignant" and "Primary lung cancer" were considered as positive, whereas "benign" and "Benign nodule" were treated as negative. A fifth label called "Suspicious malignant nodule" appeared two times. Since the diagnosis was not clearly defined for those images, they were not included in the training data to avoid any noise.

\subsubsection{From DICOM to augmented NumPy arrays}
\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}Like PROSTATEx, the processing was split into two parts: preprocessing (DICOM to NumPy arrays) and augmentation (NumPy arrays to cropped augmented NumPy arrays). Algorithm \ref{alg:LungCTChallenge_preprocessing} shows the various preprocessing steps.

The augmentation process is close to the one applied to PROSTATEx. An advantage Lung CT Challenge has is that all images have the same resolution (512x512 pixels), making the alignment process used for PROSTATEx useless. Nevertheless, lung tumors are particular. What differentiates them from the others is their look. In fact, lung tumors usually look like spider webs with a central nodule and branches coming out of it, whereas brain and prostate tumors tend to have well-defined boundaries. Therefore, a larger patch size was chosen in order to keep the extra information provided by the branches in the peripheral tissue. Furthermore, a perfect cropping method would only crop the interesting part of the tissue and a margin around it. However, tumor size varies a lot. It can either be extremely small (a dot) or large (massive nodule with long branches). Consequently, the selected patch size makes the largest tumors fit perfectly in the cropped area, inducing more tissue for smaller tumors. Naturally, there were some outliers whose size was a lot larger than the chosen patch size. If the patch size was set accordingly, entire lungs were sometimes visible in other patients' case. For this reason, these large tumors were treated as outliers. The chosen dimensions are 1/6 of the image width and height (i.e. 85x85px since the image size is 512x512px). These patches were then resized to 65x65px thanks to a bicubic interpolation, before being exported as NumPy arrays.

@@

\subsection{Kaggle Brain}
\label{sec:kaggleBrain}
\subsubsection{Dataset description}
\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}This dataset is publicly available on Kaggle~\cite{51}. All images come from single-sequence MRIs. It is composed of two folders: one containing images with cancerous tumors and another one containing images either without tumor or with a non-cancerous tumor. Images whose quality was extremely low (resolution, picture of a screen, artefacts, etc.) were manually removed. In other cases, the image quality was good but disturbing objects such as left/right markers or arrows pointing the tumor were visible. These objects were manually removed using the object removal tool in Photoshop or by cropping the image if the disturbing object was far from the skull. At the end of this cleaning process, the dataset was composed of 96 benign lesions and 133 malignant ones. File names and file formats were still heterogeneous.

@@

The first step was to standardize the file names and formats. The chosen convention is "$[yes|no]{index}.jpg$". The vast majority of the files were in the JPG file format, so the few PNG images were converted. Then, a CSV file containing the following columns was created: "PatientID", "Nodule Center Position", "fid", "Diagnosis". The "PatientID" is the filename without extension that was set during the first step (for example "$yes99$" or "$no3$"). The "Nodule Center Position" represents the pixel coordinates of the lesion. Fortunately, brain tumors are clearly visible on brain MRIs, which eased the process. To find the coordinates, every single image was opened at its original size in an image visualization software. A rectangle selection tool was used to draw a rectangle from the top left corner of the image (because the top left pixel corresponds to the pixel at the coordinates $(0,0)$ in NumPy arrays) to the center of the lesion. The pixel width and height of this rectangle gave the "i,j" coordinates of the lesion. These coordinates were reported manually. Then, the "fid" corresponds to the Finding ID. This parameter would be more pertinent if the same "PatientID" could have multiple findings, i.e. if the same patient could have multiple tumors. Since this information is made use of during the image processing of the other datasets, it was also added here. In this particular case, the "fid" was always set to "1". Finally, the "Diagnosis" was defined as "False" or "True" according to the folder which the image was in. This allows to simply use the same processing techniques as for the other datasets.

@@

Similarly to Lung CT Challenge, tumor sizes vary substantially. The same methodology was applied regarding the patch size, i.e. choosing a size which fits the majority of large tumors perfectly. Even though the images all show the head under the same perspective (same height/width ratio, visible skull, black margin around the skull), images were probably collected from different sources, which explains the heterogeneity of resolutions. As a consequence, the patch size could not be a fixed number of pixels for all images because the cropped area would have been the entire head on one image and a small portion of the tumor on another. The trick consisted in finding the horizontal and vertical portions of the image occupied by the tumor. In fact, cropping an area centered on the lesion whose height was $\frac{2}{7}$ of the image and whose width was $\frac{1}{4}$ gave good results. Finally, images were also resized to 65x65px using a bicubic interpolation before being saved as NumPy arrays.

\subsection{Image cropping}
\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}Multiple ways of cropping images were tested out. The differences lay in how much tissue was kept around the lesion. In some cases, more context around the lesion allows to capture details about the connections between the lesion and blood vessels for example. Moreover, the abrupt borders inducing a clear change of colors can help detect tumors. Experience showed that more context helps when the images are single-sequence. On the other hand, cropping right around the lesion or even in the lesion gives better results for multi-sequence data, i.e. when T2-weighted, DWI and ADC images (MRIs) are simultaneously used to perform the tumor detection and classification. In the latter case, the lesion usually appears distinctively on the three sequences, which impacts the decision a lot. On single-sequence images, there is a clear lack of information which makes it difficult to differentiate a tumor from another tissue without much context.

\subsection{Verification}
\subsubsection{Visual checking}
\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}Regarding the Kaggle brain dataset, the data generated was paid close attention to as the ground truth CSV file was created by hand (see Section \ref{sec:braingroundtruth}). In fact, each generated image was taken a close look at, which allowed to spot 3 mistakes over the whole dataset. They concerned typos in the "Nodule Center Position", which created pitch black images since the cropped area was outside of the original images.
Like PROSTATEx (see Section \ref{sec:dataProcessingVerification}), images with red dots were also generated for the lung and brain datasets.

\section{Transfer learning implementation}
\subsection{Layer freezing}
\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}After each dataset switch, the model is first trained with freezed layers. The freezed layers are those considered as feature extractor, i.e. the first layers of the model. The exact point where the model is split is arbitrary. In our case, it is located after the convolutions. To perform this freezing, multiple steps are required:
\begin{enumerate}
	\item Reset the last layers. To achieve this, an instance of the model is created by loading the state dictionary of the previous one, before replacing the module containing the last layers with a new instance (i.e. a new initialization). 
	\item Iterate over the batch normalization layers and freeze them. Step 3 is not sufficient to freeze batch normalization layers. In an implementation based on PyTorch, this can be achieved by calling \textit{module.eval()} on each module containing such layers (after the \textit{model.train()} call). 
	\item Only pass the parameters of these last layers to the optimizer.
\end{enumerate}

@@

\subsection{Conditions to choose the best model}
\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}At each step of the transfer learning, the best model is saved in order to be used during the next transfer learning step. After each epoch, the current model is compared to the previous best model. If it is better, the current model becomes the best model. A model is considered as better if its AUC is higher than the best model's AUC, and if its accuracy is higher than the best model's accuracy. As the goal is to optimize the AUC, the first condition makes immediate sense. The second condition was put in place in order to prevent a sudden spike in AUC to set an unreacheable value. This scenario did not occur during this experiment but during previous ones with random initialization. The AUC at early epochs could be really high, saving an untrained model as best model.

@@

\subsubsection{Automated transfer learning}
\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin}}The first Python script runs the entire transfer learning pipeline by itself. Multiple command line arguments are required in order to load the right datasets and the right model, to define the hyperparameters, etc. Table \ref{table:transfer_learning_options} gives an exhaustive list of the script options. The particularity lies in the learning rate, number of epochs, batch size and dropout options. In fact, the pipeline consists in seven different phases of training: dataset 1 (full model), dataset 2 (frozen model), dataset 2 (full), dataset 3 (frozen), dataset 3 (full), dataset 4 (frozen) and dataset 4 (full). In such a case, each phase uses different hyperparameters. Therefore, seven different parameters need to be specified for these command line options. For example, a valid value for the "-{}-lr" option could be "1e-8,1e-5,1e-6,1e-6,1e-5,1e-7,1e-8".

\subsubsection{Manual transfer learning}
\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}The second Python script allows to perform each transfer learning step by hand. Thanks to this, a good combination of hyperparameters can be searched for without having to run the whole pipeline. The command line arguments slighty differ from the automated version. Table \ref{table:manual_transfer_learning_options} lists them. First of all, a single training set and a single validation set have to be specificied. In addition to this, a second dataset is required: the reference/target one, which does not change throughout the process. This dataset is useful for the global performance graphs (see Section \ref{sec:global_performance}). Another parameter used for the global performance graphs is the "-{}-last{\_}layer". Its role is limited to this, unless "-{}-{}attach{\_}DM" is set to "True". In this case, the decision maker loaded as "-{}-last{\_}layer" is attached to the current model, provided the current step is a frozen one. This option is used at the last frozen step (target dataset), as this decision maker comes from the first training (target dataset as well). Doing this instead of a using a new random decision maker (as it is the case for the non-target datasets) allows to start with a skilled decision maker and optimize it according to the features learned during the transfer learning. Furthermore, arguments which were comma-separated values become single values, as this file runs a single transfer learning step instead of seven. Also, an output name for Tensorboard needs to be specified. In the automated version, this output name is automatically handled by the code itself. The latter allows to organize the various graphs generated by Tensorboard within the Tensorboard interface. If it is used in a correct way, Tensorboard can then be run on the root directory containing all graphs (which may be in subdirectories).

@@

@@

\subsection{Visualization of the impact of the various datasets on the target task}
\label{sec:global_performance}
\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}The so-called global performance graphs are a visual representation of the impact of each part of the transfer learning and each dataset on the target task. On paper, each step of the transfer learning is supposed to make the model learn new features that are useful for the first dataset (the target dataset). As described in Section \ref{sec:process_overview}, the model can be seen as a feature extractor and a decision maker. Hence, to see how much each step is actually helping the process, the decision maker resulting from the best model of the first step of the transfer learning, i.e. the training on the first/target dataset, is saved. As this decision maker is saved after the training on the target dataset, it is specialized in the classification of the latter. In other words, if the target dataset is a prostate dataset, it will classify prostate tumors accurately.

Then, after each epoch of every other transfer learning step, the saved decision maker is attached to the current feature extractor. This temporary model is then tested on the target training and validation sets, which results in different training and validation metrics.

At the end of the transfer learning pipeline, a graph is generated for each metric. These curves show the evolution of the latter across the entire process, which allows to determine the impact of each non-target dataset on the target task. If a metric is increasing, the features learned thanks to the corresponding dataset are useful for the target one. In this case, the graphs show how much the features learned from each dataset help classify prostate lesions.

@@

\subsection{Experimental setup}
\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin}}%The whole transfer learning consists in multiple steps.
In order to maximize performance during the whole process, the hyperparameters were manually tuned at each step, following a grid search approach to find the best ones. Table \ref{table:tl_hyperparameters_at_each_step} presents the best hyperparameters used at each step of the experiment. Note that, as it was the case in the paper experiment reproduction (see Section \ref{sec:paper_reproduction_experimental_setup}), the first model was also initialized by loading the best model among 200 initializations. The model was implemented with PyTorch using the Adam optimizer to update its weights. The complete succession of the process took approximately 60 hours on an Nvidia GeForce RTX 2080 Ti. 
%Tesla v100 XSM2 graphics card.

\begin{table}[H]
\centering
\resizebox{0.9\textwidth}{!}{%
\begin{tabular}{|l|l|c|c|c|c|}
\hline
\rowcolor[HTML]{DAE8FC} 
\textbf{Dataset}         & \multicolumn{1}{c|}{\cellcolor[HTML]{DAE8FC}\textbf{Step}} & \textbf{Learning rate} & \multicolumn{1}{c|}{\cellcolor[HTML]{DAE8FC}\textbf{Batch size}} & \textbf{Dropout} & \textbf{Number of epochs} \\ \hline
\textbf{PROSTATEx}       & DS1/Full                                                   & 1e-8                   & 128                                                              & 0.4              & 2000                      \\ \hline
\textbf{Kaggle Brain}    & DS2/Frozen                                                 & 1e-7                   & 128                                                              & 0.3              & 2000                      \\ \hline
\textbf{Kaggle Brain}    & DS2/Full                                                   & 1e-8                   & 128                                                              & 0.3              & 2000                      \\ \hline
\textbf{Lung CT Challenge} & DS3/Frozen                                                 & 1e-5                   & 128                                                              & 0.3              & 2000                      \\ \hline
\textbf{Lung CT Challenge} & DS3/Full                                                   & 1e-8                   & 128                                                              & 0.3              & 2000                      \\ \hline
\textbf{PROSTATEx}       & DS4/Frozen                                                 & 1e-5                   & 128                                                              & 0.3              & 2000                      \\ \hline
\textbf{PROSTATEx}       & DS4/Full                                                   & 1e-9                   & 128                                                              & 0.0              & 2000                      \\ \hline
\end{tabular}%
}
\caption{Transfer learning -- Hyperparameters at each step}
\label{table:tl_hyperparameters_at_each_step}
\end{table}
